Biomass and species richness temporal trends drives the assembly and the disassembly of food-web in natural fish assemblages

### 2021-06-14 paper/manuscript.Rmd
begin to revise results and methods

```{r}
loadd(full_data2, st_mono_trends_combined_list)
```
```{r}
nb_st <- unique(full_data2$station) %>% length
rm(full_data2)
nb_st_var <- map_dbl(st_mono_trends_combined_list$station, length)
names(nb_st_var) <- st_mono_trends_combined_list$variable
```

@@

With the goal to characterize the links among the temporal variations of network
and community structure and the temporal variations of community biomass and
species richness, we characterized the temporal trajectories of network and
community structure at each site. We used a method based on second order
polynomial which allows to assess both the direction and the monotony of
temporal trends [@rigal_method_2020]. To avoid complex variations of biomass and
species richness, we selected the stations that displayed monotonic trends for
these variable. Over the `r nb_st` stations, the selection procedure resulted
in respectively `r nb_st_var["log_bm_std"]` stations  and `r nb_st_var["log_rich_std"]`
for community biomass and species richness analysis.

@@

We selected the same stations that was used for the temporal analysis.

@@

With the aim to test (1) the relationship among the temporal dynamics of network
structure from one side and species richness and community biomass from another
side, (2) if those relationships were different if the communities were
assembling or disassembling, (3) if those relationships were different if
small/big communities, we modelled the dynamics the temporal trends of food-web
structure ($dN$) in function of the temporal trends of community biomass ($dB$) and species
richness ($dR$), if the latter was positive of not (Inc) and with respectively the median
community biomass ($B_i$) and species richness ($R_i$)of the three first years of the temporal
series.

$dN = dB + B_i + dB \times B + Inc_B \times B_i + dB \times Inc_B \times B_i$

$dN = dR + R_i + dR \times R_i + Inc_R \times R_i + dR \times Inc_R \times R_i$

@@

  - $dN$: temporal trend of network structure and communities structure
  - $dB$: temporal trend of community (log) biomass
  - $dR$: temporal trend of (log) species richness 
  - $B_i$: median community biomass of the first three year
  - $R_i$: median species richness of the first three year
  - $Inc_B$: Is $dB > 0$?
  - $Inc_R$: Is $dR > 0$?

@@

```{r}
loadd(rsq_sp_mod, summary_table, sp_models, model_summary)
rsq <- summary_table %>%
  filter(y %in% unique(c(net_var_bm, net_var_rich))) %>%
select(y, x, mod_summary_table) %>%
unnest(mod_summary_table) %>%
select(y, x, r.squared, adj.r.squared)
```

```{r}
rsq_nind_bm <- rsq[rsq$y == "piel_nind" & rsq$x == "log_bm_std",]$adj.r.squared %>%
  round(., 3)
rsq_bm_rich <- rsq[rsq$y == "log_bm_std" & rsq$x == "log_rich_std",]$adj.r.squared %>%
  round(., 3)
rsq_tlvl_rich_sp <- rsq_sp_mod$log_rich_std[
  rsq_sp_mod$log_rich_std$Response == "w_trph_lvl_avg", 
  ]$Conditional %>%
    round(., 3)
rsq_piel_bm_rich_sp <- rsq_sp_mod$log_bm_std[
  rsq_sp_mod$log_rich_std$Response == "piel_bm", 
  ]$Conditional %>%
    round(., 3)
```

Overall, we found high heterogeneity in the explicative power of statistical
models linking temporal dynamics of food-web structure and the one of community
biomass and species richness spanning from $R_{adj}^2= `r rsq_nind_bm`$ for the
relationships between to $R_{adj}^2= `r rsq_bm_rich`$ between biomass and
species richness (Table \@ref(tab:rsq)). The spatial relationships among
food-web structure aspects were more performant, with $R_{cond}^2$ varing from 
`r rsq_tlvl_rich_sp` to `r rsq_piel_bm_rich_sp` (Table \@ref(tab:rsq-sp)).

Overall, we found no significant effect of the temporal gain or losses of
communinity biomass biomass and species richness on the relationship between the
temporal trends of community structure and that of biomass and species richness
(Table \@ref(tab:aov-bm), \@ref(tab:aov-rich)).

@@

```{r}
loadd(model_log_bm_f3y_table_medium, model_log_rich_f3y_table_medium)
reg_bm <- model_log_bm_f3y_table_medium$reg_table
#reg_bm[reg_bm$y == ]
```
We found a significant positive relationship between the temporal trends of
community biomass from one side and species richness, average trophic level,
connectance from the other side (Fig. \@ref(fig:bm), Table \@ref(tab:aov-bm)),
meaning that they temporally increased when the community biomass temporally increased and
respectively that they temporally decreased when the community biomass decreased in
time.

We found a significant negative relationship between the temporal trends of
community biomass and that of the pielou of biomass (Figure \@ref(fig:bm), Table
\@ref(tab:aov-bm)), meaning that the eveness of biomass across species
temporally increased when the community biomass temporally decreased and
respectively that the evenness temporally decreased when the community biomass
temporally increased (Fig. \@ref(fig:bm), Table \@ref(tab:aov-bm)). However, we
found no significant relationship between the temporal trends of community biomass and that
of the pielou of individual.

Overall, we did not find effect of assembly/disassembly on the relationships
between the temporal trends of community structure and the one of community
biomass. We also found that the median community biomass had no effect on the
relationship between the temporal trends of community structure and the one of community
biomass.

@@

The temporal trends of community biomass, average trophic level, pielou of biomass,
pielou of individual had a positive relationship with that of species richness
(Fig. \@ref(fig:rich), Table \@ref(tab:aov-rich)). However, the direct
relationship between the temporal trends of average trophic level and the one of
species richness was not significant after controling for community biomass temporal
trends [**TO CHECK: refaire tous les modèles temporels en mettant à la fois la
pente des biomasses et de la richness en même temps??**].

We also found that the relationship between the temporal trends of pielou of
individual and connectance and that of species richness was dependant of the
overall species richness of the community (Fig. \@ref(fig:rich), Table
\@ref(tab:aov-rich)), meaning that the relationship between temporal trends of
pielou of individual and that of species richness was more positive as the
community contained more species (Fig. \@ref(fig:rich), Table
\@ref(tab:reg-rich)). While the relationship between temporal trends of pielou
of individual and that of species richness was more negative as the community
contained more species (Fig. \@ref(fig:rich), Table \@ref(tab:reg-rich)).

@@

We found that species richness, connectance, average trophic level were
spatially positively related to community biomass while the pielou of biomass was negatively related to
community biomass (Fig. \@ref(fig:bm), Table \@ref(tab:aov-rich-sp)). However,
we found no spatial relation between the pielou of individual and community
biomass.

@@

We found that community biomass, the pielou of biomass and individual were
positively related to species richness while average trophic level and
connectance were negatively related to
community biomass (Fig. \@ref(fig:rich), Table \@ref(tab:aov-rich-sp)).

@@

```{r}
loadd(pred_bm_rich_mono_stable_trends, anova_table_bm_rich_mono_stable_trends)
anova_table_bm_rich_mono_stable_trends %>%
  mutate(
    response = str_replace_all(response, var_replacement()),
    term = str_replace_all(term, var_replacement())
  ) %>%
  kable(.,
    booktabs = TRUE,
    row.names = FALSE,
    caption =
      paste0(
      "ANOVA results for the links between the temporal trends of
      community biomass and the temporal trends of the network structure."
    ),
  label = "aov-bm-rich"
  ) %>%
  kable_styling(latex_options = c("scale_down", "stripped")) %>%
  collapse_rows(columns = 1, valign = "top")
```

```{r}
loadd(reg_table_bm_rich_mono_stable_trends)
reg_table_bm_rich_mono_stable_trends %>%
  mutate(
    response = str_replace_all(response, var_replacement()),
    term = str_replace_all(term, var_replacement())
    ) %>%
kable(., booktabs = TRUE,
  row.names = FALSE,
  caption =
    paste0(
      "Regression coefficients for the linear models explaining the links
      between the temporal trends of community biomass and the temporal trends
      of the network structure."),
      label = "reg-bm-rich"
      ) %>%
  kable_styling(latex_options = c("scale_down", "striped")) %>%
  collapse_rows(columns = 1, valign = "top")
```

```{r}
loadd(rsq_table_bm_rich_mono_stable_trends)
rsq_table_bm_rich_mono_stable_trends %>%
  mutate(
    Response = str_replace_all(Response, var_replacement())
  ) %>%
  kable(., booktabs = TRUE,
    caption =
      paste0(
      "Marginal and conditional squared R for temporal models."),
  label = "rsq-sp"
      ) %>% #
  kable_styling(latex_options = c("striped")) %>%
  collapse_rows(columns = 1, valign = "top")
```

@@

```{r}
sp_anova_table_bm_rich_mono_stable_trends %>%
  mutate(
    response = str_replace_all(response, var_replacement()),
    term = str_replace_all(term, var_replacement()),
  ) %>%
  kable(., booktabs = TRUE,
    caption =
      paste0(
      "ANOVA results for the links among the spatial variation of community
      biomass and species richness and the one of network structure."),
  label = "aov-sp"
  ) %>%
  kable_styling(latex_options = c("scale_down", "striped")) %>%
  collapse_rows(columns = 1, valign = "top")
```

>>>>

```{r}
loadd(slope_com_var_no_covar, st_mono_trends_stable_rich_bm)
```
```{r}
nb_st <- unique(slope_com_var_no_covar$station) %>% length
rm(slope_com_var_no_covar)
nb_st_rich_bm <- st_mono_trends_stable_rich_bm %>% length
```

@@

To characterize the links among the temporal trends of network and community
structure and the temporal trends of community biomass and species richness, we
characterized the temporal trajectories of network and community structure at
each site. We used a method based on second order polynomial which allows to
assess both the direction and the monotony of temporal trends
[@rigal_method_2020]. To avoid complex variations of biomass and species
richness, we selected the stations that displayed monotonic trends for these
variables, i.e. the ones for which the second order polynomial term was non
significant. Over the `r nb_st` stations, the selection procedure resulted in 
`r nb_st_rich_bm` for community biomass and species richness analysis.

There was a clear relationship between the size of the communities and their
temporal dynamics. Communities that displayed low community biomass at the
beginning of their monitoring tend to display increase in biomass through time
and reversely. There same the same pattern for species richness, with
communities that displayed high species richness at the
beginning of their monitoring tend to display decrease in species richness through time
and reversely.

@@

In spatial analysis, we selected the same stations that were used for the
temporal analysis. We summarised temporal series of each stations by the median
value over time of each variable.

@@

With the aim to test the relationship among the temporal dynamics of network
structure from one side and species richness and community biomass from another
side, we modelled the dynamics the temporal trends of food-web
structure ($dN$) in function of the temporal trends of community biomass ($dB$) and species
richness ($dR$).

$dN = dB + dR$

@@

  - $dN$: temporal trend of network structure and communities structure
  - $dB$: temporal trend of community (log) biomass
  - $dR$: temporal trend of (log) species richness

@@

```{r}
loadd(rsq_sp_mod, summary_table, sp_models, model_summary)

loadd(rsq_table_bm_rich_mono_stable_trends, sp_rsq_table_bm_rich_mono_stable_trends)
```

```{r}
rsq_piel_nind <- rsq_table_bm_rich_mono_stable_trends %>%
  filter(Response == "piel_nind") %>%
  .$R.squared %>%
  round(., 3)

rsq_tlvl <- rsq_table_bm_rich_mono_stable_trends %>%
  filter(Response == "w_trph_lvl_avg") %>%
  .$R.squared %>%
  round(., 3)

rsq_tlvl_sp <- sp_rsq_table_bm_rich_mono_stable_trends %>%
  filter(Response == "w_trph_lvl_avg") %>%
  .$Conditional %>%
  round(., 3)

rsq_piel_bm_sp <- sp_rsq_table_bm_rich_mono_stable_trends %>%
  filter(Response == "piel_bm") %>%
  .$Conditional %>%
  round(., 3)
```

Overall, we found high heterogeneity in the explicative power of statistical
models linking temporal dynamics of food-web structure and the one of community
biomass and species richness spanning from $R_{adj}^2= `r rsq_piel_nind`$ for
pielou of individual number to $R_{adj}^2= `r rsq_tlvl`$ for average trophic
level (Table \@ref(tab:rsq)). The spatial models
 were more performant, with $R_{cond}^2$ varing from 
`r rsq_tlvl_sp` to `r rsq_piel_bm_sp` (Table \@ref(tab:rsq-sp)).

---
#Overall, we found no significant effect of the temporal gain or losses of
#communinity biomass biomass and species richness on the relationship between the
#temporal trends of community structure and that of biomass and species richness
#(Table \@ref(tab:aov)).
---

@@

```{r}
loadd(model_log_bm_f3y_table_medium, model_log_rich_f3y_table_medium)
reg_bm <- model_log_bm_f3y_table_medium$reg_table
#reg_bm[reg_bm$y == ]
```
We found a significant positive relationship between the temporal trends of
community biomass from one side and species richness, average trophic level,
connectance from the other side (Fig. \@ref(fig:bm), Table \@ref(tab:aov)),
meaning that they temporally increased when the community biomass temporally increased and
respectively that they temporally decreased when the community biomass decreased in
time.

We found a significant negative relationship between the temporal trends of
community biomass and that of the pielou of biomass (Figure \@ref(fig:bm), Table
\@ref(tab:aov)), meaning that the eveness of biomass across species
temporally increased when the community biomass temporally decreased and
respectively that the evenness temporally decreased when the community biomass
temporally increased (Fig. \@ref(fig:bm), Table \@ref(tab:aov)). We observed the
same pattern between the temporal trends of
community biomass and that of the pielou of individual number.

---
#Overall, we did not find effect of assembly/disassembly on the relationships
#between the temporal trends of community structure and the one of community
#biomass. We also found that the median community biomass had no effect on the
#relationship between the temporal trends of community structure and the one of community
#biomass.
---

@@

HERE
The temporal trends of community biomass, average trophic level, pielou of biomass,
pielou of individual had a positive relationship with that of species richness
(Fig. \@ref(fig:rich), Table \@ref(tab:aov)). However, the direct
relationship between the temporal trends of average trophic level and the one of
species richness was not significant after controling for community biomass temporal
trends [**TO CHECK: refaire tous les modèles temporels en mettant à la fois la
pente des biomasses et de la richness en même temps??**].

We also found that the relationship between the temporal trends of pielou of
individual and connectance and that of species richness was dependant of the
overall species richness of the community (Fig. \@ref(fig:rich), Table
\@ref(tab:aov)), meaning that the relationship between temporal trends of
pielou of individual and that of species richness was more positive as the
community contained more species (Fig. \@ref(fig:rich), Table
\@ref(tab:reg)). While the relationship between temporal trends of pielou
of individual and that of species richness was more negative as the community
contained more species (Fig. \@ref(fig:rich), Table \@ref(tab:reg)).

@@

We found that species richness, connectance, average trophic level were
spatially positively related to community biomass while the pielou of biomass was negatively related to
community biomass (Fig. \@ref(fig:bm), Table \@ref(tab:aov-sp)). However,
we found no spatial relation between the pielou of individual and community
biomass.

@@

We found that community biomass, the pielou of biomass and individual were
positively related to species richness while average trophic level and
connectance were negatively related to
community biomass (Fig. \@ref(fig:rich), Table \@ref(tab:aov-sp)).

@@

```{r}
loadd(pred_bm_rich_mono_stable_trends, anova_table_bm_rich_mono_stable_trends)
anova_table_bm_rich_mono_stable_trends %>%
  mutate(
    response = str_replace_all(response, var_replacement()),
    term = str_replace_all(term, var_replacement())
  ) %>%
  kable(.,
    booktabs = TRUE,
    row.names = FALSE,
    caption =
      paste0(
      "ANOVA results for the links between the temporal trends of
      community biomass and the temporal trends of the network structure."
    ),
  label = "aov"
  ) %>%
  kable_styling(latex_options = c("scale_down", "stripped")) %>%
  collapse_rows(columns = 1, valign = "top")
```

```{r}
loadd(reg_table_bm_rich_mono_stable_trends)
reg_table_bm_rich_mono_stable_trends %>%
  mutate(
    response = str_replace_all(response, var_replacement()),
    term = str_replace_all(term, var_replacement())
    ) %>%
kable(., booktabs = TRUE,
  row.names = FALSE,
  caption =
    paste0(
      "Regression coefficients for the linear models explaining the links
      between the temporal trends of community biomass and the temporal trends
      of the network structure."),
      label = "reg"
      ) %>%
  kable_styling(latex_options = c("scale_down", "striped")) %>%
  collapse_rows(columns = 1, valign = "top")
```

```{r}
loadd(rsq_table_bm_rich_mono_stable_trends)
rsq_table_bm_rich_mono_stable_trends %>%
  mutate(
    Response = str_replace_all(Response, var_replacement())
  ) %>%
  kable(., booktabs = TRUE,
    caption =
      paste0(
      "Marginal and conditional squared R for temporal models."),
  label = "rsq"
      ) %>% #
  kable_styling(latex_options = c("striped")) %>%
  collapse_rows(columns = 1, valign = "top")
```

@@

```{r}
sp_anova_table_bm_rich_mono_stable_trends %>%
  mutate(
    response = str_replace_all(response, var_replacement()),
    term = str_replace_all(term, var_replacement()),
  ) %>%
  kable(.,
    booktabs = TRUE,
    row.names = FALSE,
    caption =
      paste0(
      "ANOVA results for the links among the spatial variation of community
      biomass and species richness and the one of network structure."),
  label = "aov-sp"
  ) %>%
  kable_styling(latex_options = c("scale_down", "striped")) %>%
  collapse_rows(columns = 1, valign = "top")
```
